---
title: "Interim Presentation"
author: "Dylan Sun & Zijiang Yang"
date: '2017-01-17'
output:
  slidy_presentation: 
    font_adjustment: +2
---

```{r, message=FALSE, warning=FALSE, echo = FALSE, results = "hide"}
library(readxl)
library(knitr)
library(data.table)
library(corrplot)
library(broom)
library(stargazer)
library(ggplot2)
library(gridExtra)
prostate <- read_excel("~/Downloads/Prostate SBRT Sexual Function Data.xlsx", skip = 7)
prostate <- data.table(prostate)
prostate <- prostate[complete.cases(prostate)] # remove last row which is empty
# Code ADT as numeric; Y = 1, N = 0
prostate[ADT == "Y", ADT := 1]
prostate[ADT == "N", ADT := 0]
prostate[, ADT := as.numeric(ADT)]
prostate[, centered_age := Age - mean(Age)]
```


## Objectives
- Determining the erectile function rates following SBRT
- Determining predictors of erectile function preservation after SBRT
- If you undergo SBRT, how likely is it that you will lose erectile function?

## Potential Predictors
Negatively Associated:
- Age, Gleason Score, T-stage, PSA, ADT, BMI

Positively Associated:
- HRQOL, Erectile Function at Baseline

We expect HRQOL and Erectile Function at Baseline to be highly correlated, because erectile function is included in the questionnaire. 


## How many people lose function?
- At baseline:
158 with function, 167 without (0.49)

- At two years:
102 with function, 223 without (0.31)

- A seemingly large number of people lose function in just two years

```{r echo=FALSE}
nf_nf <- dim(prostate[`Erectile Function at Baseline` == 0 & `Erectile Function at 2 years` == 0])[1]
nf_hf <- dim(prostate[`Erectile Function at Baseline` == 0 & `Erectile Function at 2 years` == 1])[1]
hf_nf <- dim(prostate[`Erectile Function at Baseline` == 1 & `Erectile Function at 2 years` == 0])[1]
hf_hf <- dim(prostate[`Erectile Function at Baseline` == 1 & `Erectile Function at 2 years` == 1])[1]
counts <- data.table(never_functional = nf_nf, gain_function = nf_hf, loss_function = hf_nf, retain_function = hf_hf)
kable(counts)
```

## Questions and Concerns
- BMI Outlier
- Subject 67 has a BMI of 112. 
This is theoretically possible.
```{r echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
bmi_hist <- ggplot(data = prostate, aes(x = BMI)) + 
  geom_histogram()
bmi_hist
```


## Multicollinearity
```{r, echo=FALSE}
correlations <- cor(prostate[, .(Age, `Gleason Score`, `T-Stage Group`, PSA, HRQOL, ADT, BMI, Erec_base = `Erectile Function at Baseline`, Erec_two = `Erectile Function at 2 years`)])
corrplot(correlations, method = "circle", tl.cex = 1.5, number.cex = 0.6)
```


## Model
$logit(EF2) = \beta_{intercept} + \beta_{age} Age + \beta_{Gleason} Gleason \\
+ \beta_{Tstage} TStage + \beta_{PSA} PSA + \beta_{HRQOL} HRQOL + \\ 
\beta_{ADT} ADT + \beta_{BMI} BMI + \beta_{EFbase} EFBase$

## Results
```{r echo=FALSE}
prostate_logreg <- glm(`Erectile Function at 2 years` ~ centered_age + `Gleason Score` + `T-Stage Group` + PSA + HRQOL + ADT + BMI + `Erectile Function at Baseline`, data = prostate, family = binomial(link = "logit"))
# summary(prostate_logreg)

# edit results table
prostate_logreg_results <- data.table(tidy(prostate_logreg))
prostate_logreg_results <- prostate_logreg_results[, lower := exp(estimate - std.error)]
prostate_logreg_results <- prostate_logreg_results[, OR := exp(estimate)]
prostate_logreg_results <- prostate_logreg_results[, upper := exp(estimate + std.error)]
kable(prostate_logreg_results[, .(term, lower, OR, upper, p.value)], digits = 2)
```

## Results
Error rate = $\dfrac{\sum |predicted-actual|}{total}$
```{r echo=FALSE}
prostate <- prostate[, predicted_prob := predict(prostate_logreg, type = "response")]
prostate <- prostate[, predicted := ifelse(predicted_prob >= 0.5, 1, 0)]
error_rate <- prostate[, sum(abs(predicted - `Erectile Function at 2 years`))]/dim(prostate)[1]
error_rate
```

## Conclusions
- Age, HRQOL, and erectile function are predictive of erectile function at two years
- PSA is somewhat protective
- Holding everything else constant, each one-year increase in baseline age increases the odds of losing erectile function by 6%. 
- Holding everything else constant, HRQOL score is protective of erectile function. 

## Prediction
```{r, echo = FALSE}
example_prostate <- data.table(centered_age = 0, "Gleason Score" =  7, "T-Stage Group" = 0, "PSA" = 7.7, 
                               HRQOL = 56, ADT = 0, BMI = 29, "Erectile Function at Baseline" =  1)
example_prostate <- example_prostate[, probability_of_function := predict(prostate_logreg, example_prostate, type = "response")]
example_prostate <- example_prostate[, predicted_function := ifelse(probability_of_function >= 0.5, 1, 0)]
example_prostate <- rbind(example_prostate, prostate[c(143, 179, 199, 231, 149), .(centered_age, `Gleason Score`, `T-Stage Group`, PSA, HRQOL, ADT, BMI, `Erectile Function at Baseline`, probability_of_function = predicted_prob, predicted_function = predicted)])[, .(cen_age = centered_age, `Gleason Score`, T_Stage = `T-Stage Group`, PSA, HRQOL, ADT, BMI, Erec_base = `Erectile Function at Baseline`, prob_func = probability_of_function, pred_func = predicted_function)]
kable(example_prostate, digits = 2)
```



